---
title: "Gold Price Predictor"
author: "Cesar Robles"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  ioslides_presentation:
    self_contained: true
  slidy_presentation: default
---
## Introduction
### **Gold Price Predictor**  
An interactive Shiny application that forecasts the average annual gold price (USD / oz) using a simple linear regression on historical GLD ETF data.  

**Key features:**  
- Input any year from **2000** up to **current**  
- Get an immediate **predicted price**  
- Explore an **interactive Plotly chart** of past data and your prediction  

## Problem Statement
Gold prices are **highly volatile** from year to year, making it hard for investors and analysts to:

- **Plan budgets** when input costs depend on gold  
- **Manage risk** in gold‐backed portfolios  
- **Forecast returns** in commodities markets  

## Problem Statement (Cont.)
Below is the recent short‐term trend, illustrating the day‑to‑day swings in the GLD ETF:
```{r plot_problem, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
library(quantmod)
library(ggplot2)

# Fetch last 20 trading days of GLD ETF
gld_xts <- getSymbols("GLD", src="yahoo", auto.assign = FALSE)
gld_df <- data.frame(
  date  = index(tail(gld_xts, 20)),
  close = as.numeric(Cl(tail(gld_xts, 20)))
)

# Mini time-series of recent GLD closes
ggplot(gld_df, aes(x = date, y = close)) +
  geom_line(color = "goldenrod", size = 1) +
  geom_point(color = "darkgoldenrod", size = 2) +
  labs(
    title = "GLD ETF: Last 20 Trading Days",
    x     = "Date",
    y     = "Close Price (USD)"
  ) +
  theme_minimal()
```

## Our Solution

- **User Input:** Select any year from 2000 up to today via a simple numeric control.  
- **Model Logic:** A linear regression trained on historical GLD ETF annual averages.  
- **Reactive Prediction:** The server instantly computes the forecasted gold price for your chosen year.  
- **Interactive Display:** An embedded Plotly chart shows the full history and highlights your predicted point.  

## Live Demo
Use the slider (or input box) below to select any year. The app will instantly predict the average gold price for that year and highlight it on the chart.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(quantmod)
library(plotly)
library(dplyr)

# Fetch & summarise annual GLD ETF data
gld_xts <- getSymbols("GLD", src="yahoo", auto.assign=FALSE)
df <- data.frame(
  date  = index(gld_xts),
  price = as.numeric(Cl(gld_xts))
) %>%
  mutate(year = as.integer(format(date, "%Y"))) %>%
  group_by(year) %>%
  summarize(price = mean(price, na.rm = TRUE)) %>%
  ungroup()

# Fit linear model
model <- lm(price ~ year, data = df)

# Build interactive Plotly figure
fig <- plot_ly(
  df, x = ~year, y = ~price,
  type = "scatter", mode = "markers",
  marker = list(color = "goldenrod", size = 6)
) %>%
  add_lines(
    x = ~year, y = fitted(model),
    line = list(color = "darkgoldenrod", width = 2)
  ) %>%
  add_markers(
    x = max(df$year), 
    y = predict(model, data.frame(year = max(df$year))),
    marker = list(color = "red", size = 10)
  ) %>%
  layout(
    title = "Gold Price Prediction",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Average GLD Close (USD)")
  )
fig
```

## Next Steps
- Explore non-linear models (polynomial regression, ARIMA)  
- Incorporate macroeconomic indicators (inflation, interest rates)  
- Gather user feedback and refine the app